Supplementary Materials

### 2023-06-28 manuscript/supplementary.tex
supp cleanup

@@

\begin{align}
  x_1 = \frac{1}{2N} \sum_{i=1}^N \left( x_{0,i}k_{0,i} + \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \right)
\end{align}

where $k_{0,i}$ is the number of surviving gametes parent $i$ passes on, and
$\delta_{0,i,j}$ is a random term that encapsulates the noise due to Mendelian
segregation of heterozygotes. If parent $i$ is a homozygote ($x_{0,i} \in \{0,
1\}$), then $\delta_{0,i,j} = 0$, whereas if $x_{0,i} = \nicefrac{1}{2}$ then
$\delta_{0,i,j} = \pm \nicefrac{1}{2}$ with equal probability. This is because
a heterozygous parent will transmit half a neutral allele in expectation, but
each round of Mendelian segregation must pass on either 0 or 1 alleles, which
the random $\pm \nicefrac{1}{2}$ term imposes. The factor of $\nicefrac{1}{2}$
is due to the fact that we're summing over $N$ diploids, but considering the
number of gametes they transmit. Since each diploid parent must have two
offspring to maintain a constant population size, $\nicefrac{1}{N} \sum_i
k_{0,i} = 2$.

@@

\begin{align}
  \Delta x_1 = x_1 - p_0 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (k_{0,i} - 1) + \frac{1}{2N} \sum_{i=1}^N \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \nonumber \\
  \Delta x_1 &= K_1 + H_1
\end{align}

where $K_1$ and $H_1$ are the random change in neutral allele frequency change
due to offspring number (including heritable and non-heritable components), and
Mendelian segregation in generation 1.

@@

\begin{align}
  \V(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \V(x_{0,i}) \V(k_{0,i} - 1) \nonumber  \\
            &\approx \frac{1}{4N} \V(x_{0,i}) \V(k_{0,i}).
\end{align}

Assuming no correlation between parental gametes (e.g. no inbreeding),
$\V(x_{0,i})$ is the binomial variance in individual allele frequency, or
$p_0(1-p_0)/2$, and $\V(k) := \V(k_{0,i})$ is the offspring variance of an
individual given by the reproduction process. For example, if the reproduction
process is a neutral multinomial Wright--Fisher, $\V(k) \approx 2$. Then, the
variance in allele frequency change due to non-heritable offspring number
variation is,

@@

\begin{align}
  \V(H_1) &= \frac{p_0(1-p_0)}{2N} \frac{1}{2}.
\end{align}

Finally, we have

@@

\begin{align}
  V_\text{WF} &:= \frac{p_0(1-p_0)}{2N_e} \nonumber \\
  V_\text{WF} &= \V(\Delta p_1) \nonumber \\
  N_e &= \frac{4N}{\V(k) + 2}
\end{align}
%
(c.f. \cite{Wright1938-tv}). Now, we look at the variance in allele frequency
change across two generations for a neutral allele. The change in neutral
allele frequency, regardless of the selective system, is directionless in
expectation. This is because whichever allele we track is arbitrary by the
definition of neutrality, which imposes a symmetry, $\E(\Delta x) = 0$. Then, the
variance of change in the first generation is,

@@

Consequently, the variance of the neutral allele frequency changes each
generation according to the probability of failing to coalescence each
generation and the pairwise covariance terms $\mathcal{C}_{i,j}$ that build up
due to associations with heritable fitness backgrounds. Note that the
covariance term $\mathcal{C}_{1,2} = \E_1\left[\E_2[\Delta x_2 \Delta x_1 |
x_1]\right] = 0$ under neutral evolution, and this simplifies to the well-known
equation for variance in a Wright--Fisher population.

Now, we turn our attention to the case where there is heritable fitness
variation, which leads to non-zero covariance terms $\mathcal{C}_{i,j}$. These
terms emerge when $\V(k)$ has a heritable component of fitness that can be
transmitted along with the neutral allele, thereby affecting the neutral
allele's trajectory in later generations. If we partition the offspring number
into heritable and non-heritable components, $k_i = 2f_i + \Vepsilon_i$, then
$\V(k_i) = 4 V_h + V_n$. When there is heritable fitness variation across
individuals, $V_h > 0$. Because the population size is assumed to be constant,
the mean heritable fitness across individuals is constrained to be $\E_i(f_i) =
1$; this implies that the individual fitness values $f_i$ are relative
fitnesses as is standard in population genetics, in this case to the population
mean. Thus $V_h$ is the squared coefficient of heritable fitness variation
(this is often denoted as $C^2 = \nicefrac{V_A}{\bar{w}^2}$, see
\cite{Crow1958-pc,Charlesworth1987-ab,Houle1992-ur}).

@@

\begin{align}
  \V(\Delta x_1) &= \V(S_1) + \V(D_1) + \V(H_1) \nonumber \\
                   &\approx \frac{p_0(1-p_0)}{2N} V_h + \frac{p_0(1-p_0)}{2N} \frac{V_n}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2} \label{supp-eq:varterms} 
\end{align}
%
(c.f. \cite{Santiago1995-hx} equation 11).

@@

After meiosis, a fraction $1-r$ of these associations between the tracked
neutral neutral allele $x_{0,1}'$ and fitness background $f_{0,i}'$ persist,
and a fraction $r$ disassociate its currently coupled background and create an
association with the homologous fitness background, $f_{0,i}''$. When linkage
is tight, this latter term can be ignored as we will do here, though it does
impact background levels of neutral diversity \parencite{Santiago1995-hx}.
Additionally, the fitness effects of the selected background may change in
later generations, in complex ways \parencite{Barton1986-yh,Turelli1990-kd}.
\textcite{Santiago1995-hx} assume an equilibrium level of fitness variation
$V_h$, with the fitness variance associated with a particular background
decaying at a simple geometric decay at rate $1-\kappa$ each generation, which
is sufficient for background selection. An approximate rate of the decay can be
worked out for the particular selective system. In general, this can be a much
more complicated function; \textcite{Buffalo2019-qs} show that even fluctuating
selection can be accommodated.

@@

With these covariance terms, we now have a full model for $\V(p_t - p_0)$. We
will look at the case for $t=2$. As shown in equation \eqref{supp-eq:varterms},
the second moment of every term is a function of the variance in neutral allele
frequency in that generation, $x_t(1-x_t)$. This variance does not stay its
initial value of $x_0(1-x_0)$; it decays each generation due to coalescence
from the drift and selection processes modeled by this approach. The rate at
which this decay happens is the effective population size each generation
implied by this model. We can write,

\begin{align}
    \label{eq:vardecay}
  \V(x_2 - p_0) &= \E({S_1}^2) + \E(S_1 S_2) + \E({S_2}^2) \nonumber \\
                  & + \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) \nonumber \\
  \frac{\V(x_2 - p_0)}{p_0(1-p_0)} &= \frac{V_h}{2N_{e,1}} + \frac{V_h Q_2^2}{2N_{e,2}}\left(1-\frac{1}{2N_{e,1}}\right)  +  \frac{V_h}{2N_{e,2}}  \left(1-\frac{1}{2N_{e,1}}\right)  \nonumber \\ 
                                     &+ \frac{V_n}{8N_{e,1}} + \frac{1}{4N_{e,1}} + \frac{V_n}{8N_{e,2}}  \left(1 - \frac{1}{2N_{e,1}}\right) 
+ \frac{1}{4N_{e,2}} \left(1 - \frac{1}{2N_{e,1}}\right).
\end{align}
%
In general, the variance in allele frequency change after $t$ generations in a
system with heritable fitness due to a single locus $r$ recombination fraction
away from the neutral site is given by,

\begin{align}
  \label{eqn:var_xt}
  \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^t \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\prod_{j=1}^{i-1} \left(1-\frac{1}{2 N_{e,j}}\right).
\end{align}

We will see later that with the correct form for $Q_t^2$, this equation holds
for polygenic selection (see Supplementary Section \ref{supp:ne-polygenic}).
Additionally, we note that since the probability of coalescence (or,
equivalently, identity by descent) is proportional to the variance in allele
frequency change \parencite{Barton2000-zg}, this equation for the variance
encodes the pairwise coalescent rate of the population through time under weak
draft in the $N_{e,t}$ terms. Others have found that the genealogies under
purifying selection can be characterized by such a time-dependent effective
population size \parencite{Nicolaisen2013-gv}.

Now, we consider how to turn this expression for the variance in neutral allele
frequency at time $t$ under weak draft into a \emph{draft-effective} population
size. In an ideal Wright--Fisher population, the variance after $t$ generations
is

\begin{align}
    \label{supp-eq:drift-vt}
    \V(x_t) &= p_0(1-p_0) \left[1 - \left(1 - \frac{1}{2N_e}\right)^t \right].
\end{align}

To find an analogous expression for the asymptotic \emph{draft-effective}
population size $N_d$, we let $N_e = N_d$ in the above equation, and consider
the change in variance between generations $t$ and $t-1$, $\V(x_t) -
\V(x_{t-1})$. Using Supplementary Equation \eqref{supp-eq:drift-vt}, we have,

\begin{align}
    \V(x_t) - \V(x_{t-1}) = p_0(1-p_0) \left[ \left(1 - \frac{1}{2N_e}\right)^{t-1} - \left(1 - \frac{1}{2N_e}\right)^t  \right]
\end{align}

\textcite{Santiago1995-hx} assume the change in population size per generation
is small enough that the coalescence term each generation is constant, i.e. $(1
- \nicefrac{1}{2N_{e,t}}) \approx (1 - \nicefrac{1}{2N_e})$ for each $t$.

\begin{align}
    \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^t \left(\frac{V_h Q_i^2}{2 N_e} + \frac{V_n}{8N_e} + \frac{1}{4N_e} \right)\left(1-\frac{1}{2 N_{e}}\right)^{i-1}.
\end{align}

One

Then, imagine the process burns in for $t$ generations and asking what the
equivalent

\begin{align}
    \V(x_t) &= \V(x_{t-1})

  p_0(1-p_0) \left[1 - \left(1 - \frac{1}{2N_d}\right)^t \right].
\end{align}

\begin{align}
  \label{eqn:ne_var}
  N_{d,t} = \frac{p_0(1-p_0) - \V(x_{t-1})}{2(\V(x_t) - \V(x_{t-1}))}
\end{align}
%
(cf. p. 1018, \cite{Santiago1995-hx}).

XXX here

\textcite{Santiago1995-hx} assume the change in population size per generation
is small enough that the coalescence term each generation is constant, i.e. $(1
- \nicefrac{1}{2N_{e,t}}) \approx (1 - \nicefrac{1}{2N_e})$ for each $t$.

\begin{align}
    \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^t \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\left(1-\frac{1}{2 N_{e}}\right)^{i-1}.
\end{align}

\begin{align}
  \label{eq:var_asymp}
  \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \frac{4 V_h Q_\infty^2 +V_n+2}{4 (2 N-1)}.
\end{align}

Note here, to differentiate from the drift-effective population size $N_e$ due
under non-selective causes, we use $N_d$ as the \emph{draft-effective}
population size. Then, if we look at the change in the variance in allele
frequency change between adjacent generations $t$ and $t-1$, we get an implied
$N_{d,t}$

---
Solving this for equation \eqref{eq:var_asymp} gives the asymptotic effective
population size,

\begin{align}
  %N_e &= \frac{2 (2 N-1)}{4 V_h Q_\infty^2 +V_n+2} \\
  N_d &\approx \frac{4N}{4 V_h Q_\infty^2 +V_n+2}
\end{align}
%
(cf. equation 18, \cite{Santiago1995-hx}). Note if $V_h = 0$, this equation
reduces to Wright's (\citeyear{Wright1938-tv}) equation.

Levels of pairwise diversity, however, depend on the full distribution of
pairwise coalescence times, or the $N_{d,t}$. While this would require solving the
recursion defined by equations \eqref{eqn:var_xt} and \eqref{eqn:ne_var},
\textcite{Santiago1998-bs} show that a reasonably good approximation follows
from using a single effective population size for all timepoints, $N_{d,t} =
N_d$, but use the time-indexed $Q_t$. This leads to,

\begin{align}
  \label{eq:ne_t}
  N_{d,t} &\approx \frac{4N}{4 V_h Q_t^2 +V_n+2}
\end{align}
%
(c.f. equation 15 of \cite{Santiago1995-hx} and p. 1270 of \cite{Santiago2016-mu}).

\subsection{Effective Population Size Under Polygenic Selection}
\label{supp:ne-polygenic}

The results of the previous section assume only a single locus a recombination
fraction $r$ apart from the neutral locus is determining the coalescence rate. The
model of \textcite{Santiago1998-bs} extends this to the case where fitness is
polygenic. As they show in their original paper, this system can accommodate a
variety of different equilibrium selection systems as long as expressions for
$V_h$ and $1-\kappa$ can be worked out. Other authors have extended similar models to
more elaborate bbreeding structures \parencite{Wray1990-zf,Woolliams1993-qo}.

First, we extend equation \eqref{eq:ne_t} to a polygenic system. Throughout, a
multiplicative fitness model is assumed (i.e. independent effects, no
epistasis), where the fitness of individual $i$ is the product of their fitness
contributions from each locus $l$, $w_{i,l}$, giving $w_i = \prod_{l=1}^n
w_{i,l}$. Then, assuming mean fitness is one and independence between sites,
the total fitness variation is multiplicative across the locus-level fitnesses,

\begin{align}
  V_h %&= \E_i\left( w_i^2 \right) - 1 \\
      %&= \prod_{l=1}^n (1 + \alpha_{l}^2 \V(g_l))  - 1  \\
      &= \prod_{l=1}^n (1 + V_{h,l})  - 1
\end{align}

Similarly, the $Q_t^2 V_h$ term in the single-locus model is multiplicative
under the polygenic model, as the change in site-specific fitness variances are
assumed independent,

\begin{align}
  Q_t^2 V_h &= \prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)  - 1
\end{align}
%
where $Q_{t,l}$ is the cumulative impact of selection due to the selected site
$l$ (c.f. equation 4 \cite{Santiago1998-bs}). The factor of $\nicefrac{1}{2}$
is due to the fact that we are ignoring the chance that the fitness background
on the homologous chromosome recombines onto the haplotype with the tracked
neutral allele (i.e. due to the $Q_t''$ associations). In other words, only
half the fitness variation in a diploid can stay associated with the neutral
allele through the generations.

@@

\begin{align}
  N_{d,t} &\approx \frac{N}{V_h Q_t^2 + 1} \nonumber \\
          &\approx \frac{N}{\prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)} \nonumber \\
          &\approx N \exp\left(- \sum_{l=1}^n \frac{V_{h,l} Q_{t,l}^2}{2}\right)
\end{align}
%
(c.f. \cite{Santiago1998-bs} equation 4b). If we write each site $l$'s
contribution to the variance as a deviation from the average contribution,
$V_{h,l}^2 = \nicefrac{1}{n} V_h + \Vepsilon_l$, then

\begin{align}
  N_{d,t} &\approx N \exp\left(- \frac{1}{2}\sum_{l=1}^n \left(\frac{1}{n}V_{h} + \Vepsilon_l\right)  Q_{t,l}^2\right) \nonumber \\
          &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2+ \sum_{l=1}^n \Vepsilon_l  Q_{t,l}^2\right)
\end{align}
%
Santiago and Caballero assume that over evolutionary replicates, as selected
sites are randomly distributed, there is not systematic covariance between the
fitness variation at a site and the decay of its association with the neutral
site, e.g. $\E(\Vepsilon_l Q_{t,l}^2) = 0$, and thus ignore the second sum
term, leading to

@@

Our model differs from that of \textcite{Santiago1998-bs} in that we do not
integrate over the entire genome. In their 1998 model, Santiago and Caballero
consider the impact of both sites under selection on the same chromosome, as
well as sites on independently assorting chromosomes. We only consider the
contribution of a single segment under purifying selection, and the reduction
from additional segments accumulate multiplicatively (similar to our
implementation of the classic BGS model, XXX). Like their model, we imagine the
reduction experienced by a focal neutral site in the middle of the segment.
The implementation of our method differs slightly on though, see XXX.

Since $Q_{t,l}$ terms depend on $l$ only through the recombination fraction
between the neutral allele and the selected site, $r(l)$. Assuming a constant
per-basepair recombination rate of $r_\text{BP}$, $r(l) = r_\text{BP} l$ we can
approximate the sum with an integral for the asymptotic $Q_\infty$. The
reduction experienced in the middle of a chromosome is twice the reduction
experienced from each half,

\begin{align}
      N_{d,\infty} &\approx N \exp\left(- \frac{V_{h}}{2L} \sum_{l=1}^L Q_{\infty,l}^2\right) \\
              &\approx N \exp\left(- \frac{V_{h}}{M} \int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr \right) \\
\end{align}

where $M = n r_\text{BP}$ is the total segment length in Morgans. Then, the
total asymptotic effect of associations from all sites as $Q_\infty^2$.  Then,

\begin{align}
  Q_\infty^2 &= \frac{2}{M}\int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr  \\
  &=  \frac{2}{M} \int_0^{\nicefrac{M}{2}} \left(\frac{1}{1 - (1-r)Z}\right)^2 dr  \\
  &= \frac{2}{(1-Z) (2-(2-M) Z)}
\end{align}

Santiago and Caballero further approximate this; we can derive their
approximation by setting $Z = 1-\kappa$ and doing a first-order Taylor series
expansion around $\kappa$ (since the loss in variance is presumed to be small).
Then,

\begin{align}
  Q_\infty^2 = \frac{2}{M (1-Z)} + \frac{2 (M-2)}{M^2}+\frac{2 \left(M^2-4 M+4\right) (1-Z)}{M^3} + \mathcal{O}(\kappa^2)
\end{align}

Santiago and Caballero's approximation only keeps this first term, e.g. their
$Q_\infty^2 \approx 2/(M(1-Z))$, which is only accurate in the domain $M >
0.2$.

\begin{align}
  N_{d,\infty}^{\text(SC98)} \approx N \exp\left(-\frac{V_h}{(1-Z)M}\right) 
\end{align}

(c.f. \cite{Santiago1998-bs} equation 8). For small $M$, the small $\kappa$
approximation strongly deviates from the full equation; thus, throughout, we
use the unapproximated version,

>>>>

\subsection{Sources of Neutral Allele Frequency Change}

@@

\begin{align}
  x_1 = \frac{1}{2N} \sum_{i=1}^N \left( x_{0,i}k_{0,i} + \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \right)
\end{align}
%
where $k_{0,i}$ is the number of surviving gametes parent $i$ passes on, and
$\delta_{0,i,j}$ is a random term that encapsulates the noise due to Mendelian
segregation of heterozygotes. If parent $i$ is a homozygote ($x_{0,i} \in \{0,
1\}$), then $\delta_{0,i,j} = 0$, whereas if $x_{0,i} = \nicefrac{1}{2}$ then
$\delta_{0,i,j} = \pm \nicefrac{1}{2}$ with equal probability. This is because
a heterozygous parent will transmit half a neutral allele in expectation, but
each round of Mendelian segregation must pass on either 0 or 1 alleles, which
the random $\pm \nicefrac{1}{2}$ term imposes. The factor of $\nicefrac{1}{2}$
is due to the fact that we're summing over $N$ diploids, but considering the
number of gametes they transmit. Since each diploid parent must have two
offspring to maintain a constant population size, $\nicefrac{1}{N} \sum_i
k_{0,i} = 2$.

@@

\begin{align}
  \Delta x_1 = x_1 - p_0 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (k_{0,i} - 1) + \frac{1}{2N} \sum_{i=1}^N \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \nonumber \\
  \Delta x_1 &= K_1 + H_1
\end{align}
%
where $K_1$ and $H_1$ are the random change in neutral allele frequency change
due to offspring number (including heritable and non-heritable components), and
Mendelian segregation in generation 1.

\subsection{The Variance in Neutral Allele Frequency Change}

@@

\begin{align}
  \V(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \V(x_{0,i}) \V(k_{0,i} - 1) \nonumber  \\
            &\approx \frac{1}{4N} \V(x_{0,i}) \V(k_{0,i}).
\end{align}
%
Assuming no correlation between parental gametes (e.g. no inbreeding),
$\V(x_{0,i})$ is the binomial variance in individual allele frequency, or
$p_0(1-p_0)/2$, and $\V(k) := \V(k_{0,i})$ is the offspring variance of an
individual given by the reproduction process. For example, if the reproduction
process is a neutral multinomial Wright--Fisher, $\V(k) \approx 2$. Then, the
variance in allele frequency change due to non-heritable offspring number
variation is,

@@

\begin{align}
  \V(H_1) &= \frac{p_0(1-p_0)}{2N} \frac{1}{2}.
\end{align}
%
Finally, we have

@@

\begin{align}
  V_\text{WF} &:= \frac{p_0(1-p_0)}{2N_e} \nonumber \\
  V_\text{WF} &= \V(\Delta p_1) \nonumber \\
  N_e &= \frac{4N}{\V(k) + 2} \label{supp-eq:var-ne}
\end{align}
%
(c.f. \cite{Wright1938-tv}).

Now, we look at the variance in allele frequency change across two generations
for a neutral allele. The change in neutral allele frequency, regardless of the
selective system, is directionless in expectation. This is because whichever
allele we track is arbitrary by the definition of neutrality, which imposes a
symmetry, $\E(\Delta x) = 0$. Then, the variance of change in the first
generation is,

@@

Consequently, the variance of the neutral allele frequency changes each
generation according to the probability of failing to coalescence each
generation and the pairwise covariance terms $\mathcal{C}_{i,j}$ that build up
due to associations with heritable fitness backgrounds. Note that the
covariance term $\mathcal{C}_{1,2} = \E_1\left[\E_2[\Delta x_2 \Delta x_1 |
x_1]\right] = 0$ when there is no heritable fitness variation, and this
simplifies to the well-known equation for variance in a Wright--Fisher
population.

\subsection{Heritable Fitness and the Accumulation of Autocovariance}

Next, we turn our attention to the case where there is heritable fitness
variation, which leads to non-zero covariance terms $\mathcal{C}_{i,j}$. These
terms emerge when $\V(k)$ has a heritable component of fitness that can be
transmitted along with the neutral allele, thereby affecting the neutral
allele's trajectory in later generations. If we partition the offspring number
into heritable and non-heritable components, $k_i = 2f_i + \varepsilon_i$, then
$\V(k_i) = 4 V_A + V_n$. When there is heritable fitness variation across
individuals, $V_A > 0$. Because the population size is assumed to be constant,
the mean heritable fitness across individuals is constrained to be $\E_i(f_i) =
1$; this implies that the individual fitness values $f_i$ are relative
fitnesses as is standard in population genetics, in this case to the population
mean. Thus $V_A$ is the squared coefficient of heritable fitness variation
(this is often denoted as $C^2 = \nicefrac{V_A}{\bar{w}^2}$, see
\cite{Crow1958-pc,Charlesworth1987-ab,Houle1992-ur}).

@@

\begin{align}
  \V(\Delta x_1) &= \V(S_1) + \V(D_1) + \V(H_1) \nonumber \\
                   &\approx \frac{p_0(1-p_0)}{2N} V_A + \frac{p_0(1-p_0)}{2N} \frac{V_n}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2} \label{supp-eq:varterms} 
\end{align}
%
(c.f. \cite{Santiago1995-hx} equation 11).

@@

After meiosis, a fraction $1-r$ of these associations between the tracked
neutral neutral allele $x_{0,1}'$ and fitness background $f_{0,i}'$ persist,
and a fraction $r$ disassociate its currently coupled background and create an
association with the homologous fitness background, $f_{0,i}''$. When linkage
is tight, this latter term can be ignored as we will do here, though it does
impact background levels of neutral diversity \parencite{Santiago1995-hx}.
Additionally, the fitness effects of the selected background may change in
later generations, in complex ways \parencite{Barton1986-yh,Turelli1990-kd}.
\textcite{Santiago1995-hx} assume an equilibrium level of fitness variation
$V_A$, with the fitness variance associated with a particular background
decaying at a simple geometric decay at rate $1-\kappa$ each generation, which
is sufficient for background selection. An approximate rate of the decay can be
worked out for the particular selective system. In general, this can be a much
more complicated function; \textcite{Buffalo2019-qs} show that even fluctuating
selection can be accommodated.

@@

With these covariance terms, we now have a full model for $\V(p_t - p_0)$. We
will look at the case for $t=2$. As shown in Supplementary Equation \eqref{supp-eq:varterms},
the second moment of every term is a function of the variance in neutral allele
frequency in that generation, $x_t(1-x_t)$. This variance does not stay its
initial value of $x_0(1-x_0)$; it decays each generation due to coalescence
from the drift and selection processes modeled by this approach. The rate at
which this decay happens is the effective population size each generation
implied by this model. We can write,

\begin{align}
    \label{eq:vardecay}
  \V(x_2 - p_0) &= \E({S_1}^2) + \E(S_1 S_2) + \E({S_2}^2) \nonumber \\
                  & + \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) \nonumber \\
  \frac{\V(x_2 - p_0)}{p_0(1-p_0)} &= \frac{V_A}{2N} + \frac{V_A Q_2^2}{2N}\left(1-\frac{1}{2N}\right)  +  \frac{V_A}{2N_}  \left(1-\frac{1}{2N}\right)  \nonumber \\ 
                                     &+ \frac{V_n}{8N} + \frac{1}{4N} + \frac{V_n}{8N}  \left(1 - \frac{1}{2N}\right) 
+ \frac{1}{4N} \left(1 - \frac{1}{2N}\right).
\end{align}
%
In general, the variance in allele frequency change after $t$ generations in a
system with heritable fitness due to a single locus $r$ recombination fraction
away from the neutral site is given by,

\begin{align}
  \label{supp-eq:var-xt}
  \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^t \frac{1}{2 N} \left(V_A Q_i^2 + \frac{V_n}{4} + \frac{1}{2} \right) \left(1-\frac{1}{2 N}\right)^{i-1}.
\end{align}
%
We will see later that with the appropriate form for $Q_t^2$, this equation
holds for polygenic selection (see Supplementary Section
\ref{supp:ne-polygenic}). Additionally, we note that since the probability of
coalescence (or, equivalently, identity by descent) is proportional to the
variance in allele frequency change \parencite{Barton2000-zg}, this equation
for the variance encodes the pairwise coalescent rate of the population through
time under weak draft. Others have found that the genealogies under purifying
selection can be characterized by such a time-dependent effective population
size \parencite{Nicolaisen2013-gv}.

\subsection{The Draft-Effective Population Size}

Now, we consider how to turn this expression for the variance in neutral allele
frequency at time $t$ under weak draft into an expression for the
\emph{draft-effective} population size analogous to Supplementary Equation
\eqref{supp-eq:var-ne}. In an ideal Wright--Fisher population, the
instantaneous variance is $\V(\Delta x_t) =
\nicefrac{p_{t-1}(1-p_{t-1})}{2N_{d,t-1}}$. Santiago and Caballero consider the
allele frequency change at some $t$ to define their effective population size
at time $t$ (which we here are calling draft-effective population size,
$N_{d,t}$),

\begin{align}
    N_{d,t} := \frac{p_{t-1} (1-p_{t-1})}{2\V(x_t) - \V(x_{t-1})}.
\end{align}
%
The time-dependency is due to $Q_i^2$, which reflects the fact that
autocovariances build up, but at a decreasing rate. Note that levels of
pairwise diversity depend on the full distribution of pairwise coalescence
times, or the $N_{d,t}$ (e.g. see equation 4 of \cite{Santiago2016-mu} and
equation 16 of \cite{Santiago1998-bs}).

Using Supplementary Equation \eqref{supp-eq:var-xt} we have,

\begin{align}
    \label{supp-eq:var-ndt}
    N_{d,t} &\approx \frac{4N}{4 V_A Q_t^2 +V_n+2}
\end{align}
%
(cf. equation 18, \cite{Santiago1995-hx}). Note if $V_A = 0$, this equation
reduces to Wright's (\citeyear{Wright1938-tv}) equation.

In many cases, we care about the asymptotic $N_d$,

\begin{align}
    N_d := \lim_{t \to \infty} N_{d,t} \approx \frac{4N}{4 V_A Q_\infty^2 +V_n+2}.
\end{align}

We now turn to different expressions for $Q_t^2$ and $V_A$.

\subsection{The Draft-Effective Population Size Under Polygenic Selection}
\label{supp:ne-polygenic}

The results of the previous section assume only a single locus a recombination
fraction $r$ apart from the neutral locus is determining the coalescence rate. The
model of \textcite{Santiago1998-bs} extends this to the case where fitness is
polygenic. As they show in their original paper, this system can accommodate a
variety of different equilibrium selection systems as long as expressions for
$V_A$ and $1-\kappa$ can be worked out. Other authors have extended similar models to
more elaborate bbreeding structures \parencite{Wray1990-zf,Woolliams1993-qo}.

First, we extend Supplementary Equation \eqref{supp-eq:var-ndt} to a polygenic
system. Throughout, we assume a multiplicative fitness model (i.e. independent
effects, no epistasis), where the fitness of individual $i$ is the product of
their fitness contributions from each locus $l$, $w_{i,l}$, giving $w_i =
\prod_{l=1}^n w_{i,l}$. Then, assuming mean fitness is one and independence
between sites, the total fitness variation is multiplicative across the
locus-level fitnesses,

\begin{align}
  V_A %&= \E_i\left( w_i^2 \right) - 1 \\
      %&= \prod_{l=1}^n (1 + \alpha_{l}^2 \V(g_l))  - 1  \\
      &= \prod_{l=1}^n (1 + V_{h,l})  - 1
\end{align}

Similarly, the $Q_t^2 V_A$ term in the single-locus model is multiplicative
under the polygenic model, as the change in site-specific fitness variances are
assumed independent,

\begin{align}
  Q_t^2 V_A &= \prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)  - 1
\end{align}
%
where $Q_{t,l}$ is the cumulative impact of selection due to the selected site
$l$ (c.f. equation 4 \cite{Santiago1998-bs}). The factor of $\nicefrac{1}{2}$
is due to the fact that we are ignoring the chance that the fitness background
on the homologous chromosome recombines onto the haplotype with the tracked
neutral allele (i.e. due to the $Q_t''$ associations). In other words, only
half the fitness variation in a diploid can stay associated with the neutral
allele through the generations.

@@

\begin{align}
  N_{d,t} &\approx \frac{N}{V_A Q_t^2 + 1} \nonumber \\
          &\approx \frac{N}{\prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)} \nonumber \\
          &\approx N \exp\left(- \sum_{l=1}^n \frac{V_{h,l} Q_{t,l}^2}{2}\right)
\end{align}
%
(c.f. \cite{Santiago1998-bs} equation 4b). If we write each site $l$'s
contribution to the variance as a deviation from the average contribution,
$V_{h,l}^2 = \nicefrac{1}{n} V_A + \varepsilon_l$, then

\begin{align}
  N_{d,t} &\approx N \exp\left(- \frac{1}{2}\sum_{l=1}^n \left(\frac{1}{n}V_{h} + \varepsilon_l\right)  Q_{t,l}^2\right) \nonumber \\
          &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2+ \sum_{l=1}^n \varepsilon_l  Q_{t,l}^2\right)
\end{align}
%
Santiago and Caballero assume that over evolutionary replicates, as selected
sites are randomly distributed, there is not systematic covariance between the
fitness variation at a site and the decay of its association with the neutral
site, e.g. $\E(\varepsilon_l Q_{t,l}^2) = 0$, and thus ignore the second sum
term, leading to

@@

Our model differs from that of \textcite{Santiago1998-bs} in that we do not
integrate over the entire genome. In their 1998 model, Santiago and Caballero
consider the impact of both sites under selection on the same chromosome, as
well as sites on independently assorting chromosomes. We only consider the
contribution of a single segment under purifying selection, and the reduction
from additional segments accumulate multiplicatively. Like their model, we
imagine the reduction experienced by a focal neutral site in the middle of the
segment. The implementation of our method differs slightly on though, see XXX.

The $Q_{t,l}$ terms depend on $l$ only through the recombination fraction
$r(l)$ between the neutral allele and the selected site. Assuming a constant
per-basepair recombination rate of $r_\text{BP}$, $r(l) = r_\text{BP} l$ we can
approximate the sum with an integral for the asymptotic $Q_\infty$. The
reduction experienced in the middle of a chromosome is twice the reduction
experienced from each half,

\begin{align}
      N_{d,\infty} &\approx N \exp\left(- \frac{V_{h}}{2L} \sum_{l=1}^L Q_{\infty,l}^2\right) \nonumber \\
              &\approx N \exp\left(- \frac{V_{h}}{M} \int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr \right) 
\end{align}
%
where $M = n r_\text{BP}$ is the total segment length in Morgans. Then, the
total asymptotic effect of associations from all sites as $Q_\infty^2$.  Then,

\begin{align}
  Q_\infty^2 &= \frac{2}{M}\int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr  \nonumber \\
  &=  \frac{2}{M} \int_0^{\nicefrac{M}{2}} \left(\frac{1}{1 - (1-r)Z}\right)^2 dr. \nonumber  \\
  &= \frac{2}{(1-Z) (2-(2-M) Z)}
\end{align}
%
Santiago and Caballero further approximate this; we can derive their
approximation by setting $Z = 1-\kappa$ and doing a first-order Taylor series
expansion around $\kappa$ (since the loss in variance is presumed to be small).
Then,

\begin{align}
  Q_\infty^2 = \frac{2}{M (1-Z)} + \frac{2 (M-2)}{M^2}+\frac{2 \left(M^2-4 M+4\right) (1-Z)}{M^3} + \mathcal{O}(\kappa^2)
\end{align}
%
Santiago and Caballero's approximation only keeps this first term, e.g. their
$Q_\infty^2 \approx 2/(M(1-Z))$, which is only accurate in the domain $M >
0.2$.

\begin{align}
  N_{d,\infty}^{\text(SC98)} \approx N \exp\left(-\frac{V_A}{(1-Z)M}\right) 
\end{align}
%
(c.f. \cite{Santiago1998-bs} equation 8). For small $M$, the small $\kappa$
approximation strongly deviates from the full equation; thus, throughout, we
use the unapproximated version,
